---
title: "sentiment analysis"
output: html_document
---

```{r setup, include=FALSE}
library(fuzzyjoin)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

```{r}
#read in file 
sentiment <- readRDS("matches_reviews_fixed.Rds")
sentiment
```

```{r sentiment packages, cache = T}
#get bing sentiment
library(tidytext) 
bing <- get_sentiments("bing")
```

```{r}
#collapse all reviews of one restaurant into one observation 
sentimentsCollapsed <- sentiment%>%
select(business_id,text)%>%
group_by(business_id)%>%
summarize(text=paste(text,collapse=" // "))

sentimentsCollapsedExcerpt <- sentimentsCollapsed %>%
  head(1)

sentimentsCollapsedExcerpt
```


```{r}
#finds top three negative and top three positive words
goodBad <- function(text) {
words <- tibble(text) %>%
  unnest_tokens(word, text)
positive <- words %>%
  count(word, sort = T) %>%
  inner_join(bing) %>%
  filter(sentiment == "positive") %>%
  head(3)
negative <- words %>%
  count(word, sort = T) %>%
  inner_join(bing) %>%
  filter(sentiment == "negative") %>%
  head(3)
c <- rbind(positive, negative)
c
}
```


```{r}
#running on one restaurant to test
goodBad(sentimentsCollapsedExcerpt$text))
```


```{r}
#using `wordcloud` package to create wordclouds of the restaurant's reviews
library(wordcloud)
wordcloud(sentimentsCollapsedExcerpt$text, max.words = 30, scale = c(10, .1),
colors = topo.colors(n = 30), random.color = TRUE)
```


#looking at chengdu restaurant
```{r}
#filtering for chengdu only
chengdu <- sentiment %>%
  filter(name.x == "chengdu taste")
#finding top3 words
chengduWords <- goodBad(chengdu$text)

#making a bar garph
ggplot(chengduWords, aes(x = reorder(word, n), y = n, fill = sentiment)) + geom_bar(stat = "identity") + coord_flip() + xlab("Positive and Negative Words") + ylab("Count") + ggtitle("Top 3 Positive and Top 3 Negative Words")
#making a wordcloud 
wordcloud(chengdu$text, max.words = 30, scale = c(10, .1),
colors = topo.colors(n = 30), random.color = TRUE)

```
COMMENTS: We see that for Cheng Du Taste which has a grade of C, its top three positive words have more counts than its top three negative words. In it's word cloud, its hard to see any words that relate to its cleanliness.



# looking at some univariate / multivariate data
```{r, warning = F}
#more overall observations
ggplot(sentiment, aes(x = stars, fill = Inspection_Grade)) + geom_bar() + xlab("Yelp Rating") + ylab("Count of Pos and Neg Words") + ggtitle("Distribution of Yelp Ratings, filled by Inspection Grade")


ggplot(sentiment, aes(x = stars, y = Inspection_Demerits)) + geom_point() + geom_smooth(method = "lm") + xlab("Yelp Rating") + ylab("Inspection Demerits") + ggtitle("Inspection Demerits versus Yelp Rating")


```

COMMENTS: most ratings are around 3-4, lots of A grades, there does not seem to be a strong negative correlation between inspection demerits and yelp rating.

